---
title: "CE2"
author: "Fabian Krügel, Alvar Stirnimann"
date: "2024-05-03"
output: html_document
---

# Class Exercise 2

## Data Extraction

```{r, echo=FALSE}
library(rvest)
library(jsonlite)
library(httr)
library(stringr)
library(dplyr)  

# Define the base URL for the Guardian API content endpoint
url <- "https://content.guardianapis.com/search"

# Setup query parameters
api_key_data <- read.csv("api_key.csv", header = FALSE)
api_key <- api_key_data$V1[1]
base_url <- "https://content.guardianapis.com/search"
query_params <- list(
  q = "cryptocurrency AND blockchain AND bitcoin OR ethereum", # to make the search more robust and specific
  "order-by" = "newest",
  "page-size" = 200,
  "api-key" = api_key,
  "show-fields" = "body"
)

# Construct URL and make a HTTP request
url <- modify_url(base_url, query = query_params)
response <- GET(url)
```

```{r, echo=FALSE}
# Process the API response and extract article details
if (http_status(response)$category == "Success") {
  data <- content(response, "parsed")
  articles <- map(data$response$results, ~list(
    title = .x$webTitle,
    body = .x$fields$body,
    date = .x$webPublicationDate  
  ))
} else {
  stop("Failed to fetch data")
}

# Converting list to a data frame
articles_df <- map_df(data$response$results, ~data.frame(
  title = coalesce(.x$webTitle, ""),
  body = coalesce(.x$fields$body, ""),
  date = coalesce(.x$webPublicationDate, "")  
), .id = "article_id")

# Save to CSV
#write.csv(articles_df, "cryptocurrency_articles.csv", row.names = FALSE)
```


## Data Preparation

```{r, echo=FALSE}
library(tidytext)

# Combine title, body, and date into one text column for text cleaning
processed_articles_df <- articles_df %>%
  mutate(
    combined_text = paste(title, body, date),
    cleaned_text = combined_text %>%
      tolower() %>%
      str_remove_all("[[:punct:]]") %>%
      str_remove_all("[[:digit:]]")
  ) %>%
  unnest_tokens(output = word, input = cleaned_text)

# Remove stop words 
clean_tokens_df <- processed_articles_df %>%
  anti_join(stop_words, by = "word")
```


## Sentiment Analysis

```{r, echo=FALSE}
# Load the AFINN sentiment lexicon and join with tokenized words
sentiment_analysis <- clean_tokens_df %>%
  inner_join(get_sentiments("afinn"), by = "word") %>%
  group_by(article_id) %>%
  summarise(
    sentiment_score = sum(value),  
    .groups = 'drop'  
  )

# Display sentiment scores for each article, sorted by score
sentiment_analysis %>%
  arrange(desc(sentiment_score)) %>%
  print(n = Inf)  # show all 200 articles

# Show the average sentiment score
sentiment_analysis %>%
  summarise(
    average_sentiment_score = mean(sentiment_score)
  )

```
The average sentiment score is positive (19.65) meaning that articles with the keyword cryptocurrency, blockchain, bitcoin, or ethereum have a positive sentiment on average. We considered the sentiment score of each word in the articles and summed them up to get the sentiment score for each article. The newest 200 articles articles fulfilling the search criteria were analyzed (July 2021 - April 2023). More research would be necessary to assess whether the positive sentiment score is robust and whether it has changed over time.


## Visualizations

```{r, echo=FALSE}
library(ggplot2)

# Plot of sentiment analysis
sentiment_plot1 <- ggplot(sentiment_analysis, aes(x = article_id, y = sentiment_score, fill = sentiment_score > 0)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c("red", "green")) +
  labs(title = "Sentiment Analysis of Articles", x = "Article ID", y = "Sentiment Score") +
  theme_minimal()

# Add the average sentiment score to the plot
sentiment_plot2 <- ggplot(sentiment_analysis, aes(x = article_id, y = sentiment_score, fill = sentiment_score > 0)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = c("red", "green")) +
  labs(title = "Sentiment Analysis of Articles", x = "Article ID", y = "Sentiment Score") +
  theme_minimal() +
  geom_hline(yintercept = mean(sentiment_analysis$sentiment_score), color = "blue", linetype = "dashed")
```

```{r, echo=FALSE}
# Histogram of sentiment scores
sentiment_plot3 <- ggplot(sentiment_analysis, aes(x = sentiment_score)) +
  geom_histogram(bins = 30, fill = "dodgerblue", color = "black") +
  labs(title = "Distribution of Sentiment Scores Across Articles",
       x = "Sentiment Score",
       y = "Count of Articles") +
  theme_minimal()

# Save the plots
#ggsave("sentiment_plot1.png", plot = sentiment_plot1, width = 10, height = 6)
#ggsave("sentiment_plot2.png", plot = sentiment_plot2, width = 10, height = 6)
#ggsave("sentiment_plot3.png", plot = sentiment_plot3, width = 10, height = 6)

```

